---
title:: RFM for Bookbinders
output:
  html_document:
    highlight: zenburn
    theme: cosmo
    df_print: paged
    toc: yes
    code_folding: hide
    code_download: true
---

```{r r_setup, include = FALSE}
## initial settings
knitr::opts_chunk$set(
  comment = NA,
  echo = TRUE,
  error = TRUE,
  cache = FALSE,
  message = FALSE,
  dpi = 144,
  warning = FALSE
)

## width to use when printing tables etc.
options(
  width = 250,
  scipen = 100,
  max.print = 5000,
  stringsAsFactors = FALSE
)

library(radiant)
```

<style>
.btn, .form-control, pre, code, pre code {
  border-radius: 4px;
}
.table {
  width: auto;
}
ul, ol {
  padding-left: 18px;
}
code, pre, pre code {
  overflow: auto;
  white-space: pre;
  word-wrap: normal;
}
code {
  color: #c7254e;
  background-color: #f9f2f4;
}
pre {
  background-color: #ffffff;
}
</style>

Loading the data ...

```{r}
bbb <- readr::read_rds("data/bbb.rds")
register("bbb")
```

### Assessing recency

```{r fig.width = 7, fig.height = 4.67, dpi = 200}
visualize(
  bbb, 
  xvar = "last", 
  yvar = "buyer", 
  type = "bar", 
  labs = list(
    y = "Proportion of buyer = 'yes'", 
    x = "Months since last purchase (last)"
  ),
  custom = FALSE
) 
```

Creating the `recency` variable `rec_iq` using the following command in _Data > Transform_:

    rec_iq = xtile(last, 5)

```{r}
## create new variable(s)
bbb <- mutate(bbb, rec_iq = xtile(last, 5))
```

Does recency predict purchase? Are the best customers in quintile 1? The graph below shows this is indeed the case.

```{r fig.width = 7, fig.height = 4.67, dpi = 200}
visualize(
  bbb, 
  xvar = "rec_iq", 
  yvar = "buyer", 
  type = "bar", 
  labs = list(
    y = "Proportion of buyer = 'yes'", 
    x = "Recency quintiles (rec_iq)"
  ),
  custom = FALSE
) 
```

### Assessing frequency

Plots shows that purchase probility is NOT highest in the 1st quantile for frequencey (`purch`).

```{r fig.width = 7, fig.height = 4.67, dpi = 200}
visualize(
  bbb, 
  xvar = "purch", 
  yvar = "buyer", 
  type = "bar", 
  labs = list(
    y = "Proportion of buyer = 'yes'", 
    x = "Purchase frequency (purch)"
  ),
  custom = FALSE
) 
```

This means we need to 'flip' the bin numbers so the highest purchase probility is in the 1st bin (quantile). The easiest way to do this is to add `rev = TRUE` in the call to `xtile`. 

    freq_iq = xtile(purch, 5, rev = TRUE)

Alternatively, you could use:

    freq_iq = 6L - xtile(purch, 5)

```{r}
## bin variables
bbb <- mutate(bbb, freq_iq = xtile(purch, 5, rev = TRUE))
```

```{r fig.width = 7, fig.height = 4.67, dpi = 200}
visualize(
  bbb, 
  xvar = "freq_iq", 
  yvar = "buyer", 
  type = "bar", 
  labs = list(
    y = "Proportion of buyer = 'yes'", 
    x = "Frequency quintiles (freq_iq)"
  ),
  custom = FALSE
) 
```

Why are there only 4 values? Looking at the histogram below we see that the distribution of `purch` is heavily skewed (to the right). This makes it difficult to create 5 _bins_ of similar size

```{r fig.width = 7, fig.height = 4.67, dpi = 200}
visualize(bbb, xvar = "purch", color = "freq_iq")
```

### Assessing monetary value

The plot shows that purchase probility is NOT highest in the 1st quantile for `monetary` (`total`)

```{r fig.width = 7, fig.height = 4.67, dpi = 200}
visualize(
  bbb, 
  xvar = "total", 
  yvar = "buyer", 
  type = "bar", 
  labs = list(
    y = "Proportion of buyer = 'yes'", 
    x = "Monetary value (total)"
  ), 
  custom = TRUE
) + 
  theme(axis.ticks = element_blank(), axis.text.x = element_blank()) 
```

Just like we did for `frequency` we have to 'flip' quantiles so the highest purchase probility is in the 1st quantile (i.e., add `rev = TRUE`)

    mon_iq = xtile(total, 5, rev = TRUE)

```{r}
## bin variables
bbb <- mutate(bbb, mon_iq = xtile(total, 5, rev = TRUE))
```

```{r fig.width = 7, fig.height = 4.67, dpi = 200}
visualize(
  bbb, 
  xvar = "mon_iq", 
  yvar = "buyer", 
  type = "bar", 
  labs = list(
    y = "Proportion of buyer = 'yes'", 
    x = "Monetary value quantiles (mon_iq)"
  ),
  custom = FALSE
)
```

## Create the RFM index

Use _Data > Transform > Create_ to generate the RFM index

    rfm_iq = paste0(rec_iq, freq_iq, mon_iq)
    
```{r}
## create new variable(s)
bbb <- mutate(bbb, rfm_iq = paste0(rec_iq, freq_iq, mon_iq))
```

```{r fig.width = 10.5, fig.height = 4.67, dpi = 200}
visualize(
  bbb, 
  xvar = "rfm_iq", 
  yvar = "buyer", 
  type = "bar", 
  labs = list(
    y = "Proportion of buyer = 'yes'", 
    x = "Independent RFM index (rfm_iq)"
  ),
  custom = TRUE
) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_hline(yintercept = 1/9)
```

## Generate  sequential N-tile

```{r}
## create new variable(s)
bbb <- group_by(bbb, rec_iq) %>% 
  mutate(freq_sq = xtile(purch, 5, rev = TRUE)) %>% 
  ungroup()

## create new variable(s)
bbb <- group_by(bbb, rec_iq, freq_sq) %>% 
  mutate(mon_sq = xtile(total, 5, rev = TRUE)) %>% 
  ungroup()
```

```{r fig.width = 7, fig.height = 4.67, dpi = 200}
visualize(
  bbb, 
  xvar = "freq_sq", 
  yvar = "buyer", 
  type = "bar", 
  labs = list(
    y = "Proportion of buyer = 'yes'", 
    x = "Frequency quintiles (freq_sq)"
  ),
  custom = FALSE
) 
```

```{r fig.width = 7, fig.height = 4.67, dpi = 200}
visualize(
  bbb, 
  xvar = "freq_sq", 
  yvar = "buyer", 
  type = "bar", 
  labs = list(
    y = "Proportion of buyer = 'yes'", 
    x = "Frequency quintiles (freq_sq)"
  ),
  custom = FALSE
) 
```

```{r fig.width = 7, fig.height = 4.67, dpi = 200}
visualize(
  bbb, 
  xvar = "mon_sq", 
  yvar = "buyer", 
  type = "bar", 
  labs = list(
    y = "Proportion of buyer = 'yes'", 
    x = "Monetary value quintiles (mon_sq)"
  ),
  custom = FALSE
) 
```

## Generate Sequential RFM index

Use _Data > Transform > Create_ to generate the RFM index

    rfm_sq = paste0(rec_iq, freq_sq, mon_sq)
 
```{r}
## create new variable(s)
bbb <- mutate(bbb, rfm_sq = paste0(rec_iq, freq_sq, mon_sq))
```

```{r fig.width = 10.5, fig.height = 4.67, dpi = 200}
visualize(
  bbb, 
  xvar = "rfm_sq", 
  yvar = "buyer", 
  type = "bar", 
  labs = list(
    y = "Proportion of buyer = 'yes'",
    x =  "Sequential RFM index (rfm_sq)"
  ),
  custom = TRUE
) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))  +
  geom_hline(yintercept = 1/9)
```

## Resonse rate without RFM

```{r}
result <- explore(
  bbb, 
  vars = "buyer", 
  fun = c("n_obs", "mean", "min", "max")
)
summary(result, top = "fun", dec = 4)
```

## Break-even (aggr)

`r breakeven = 1 / 9`

The breakeven value is `r format_nr(breakeven, dec = 2, perc = TRUE)`. All cells above the breakeven line in the plot below will be mailed.

```{r fig.width = 10.5, fig.height = 4.67, dpi = 200}
visualize(
  bbb, 
  xvar = "rfm_iq", 
  yvar = "buyer", 
  type = "bar", 
  labs = list(
    y = "Proportion of buyer = 'yes'", 
    x = "Independent RFM index (rfm_iq)"
  ),
  custom = TRUE
) + 
  geom_hline(aes(yintercept = breakeven)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## Resonse rate

Now that we are creating variables to use in targeting we should use the `training` variable. 

Create the `mailto_iq` variable for independent RFM

    mailto_iq = mean(buyer == "yes") > 1/9

```{r}
## mail TRUE or FALSE
bbb <- group_by(bbb, rfm_iq) %>% 
  mutate(mailto_iq = (sum(training == 1 & buyer == "yes") / sum(training == 1)) > breakeven) %>% 
  ungroup()

## also calculate response rate per group as an illustration
bbb <- group_by(bbb, rfm_iq) %>% 
  mutate(rfm_iq_resp = sum(training == 1 & buyer == "yes") / sum(training == 1)) %>% 
  ungroup()
```

```{r}
result <- pivotr(
  bbb, 
  cvars = c("mailto_iq", "training"), 
  fun = "n_obs", 
  normalize = "column"
)
summary(result, dec = 2, perc = TRUE)
```

```{r}
result <- explore(
  bbb, 
  vars = "buyer", 
  byvar = c("training", "mailto_iq"), 
  fun = c("n_obs", "mean")
)
summary(result, dec = 4)
```

## Calculate profits and ROME

```{r}
dat <- filter(bbb, training == 0)
perc_mail = mean(dat$mailto_iq)
nr_mail = 500000 * perc_mail
rep_rate <- filter(dat, mailto_iq == TRUE) %>% 
  summarize(rep_rate = mean(buyer == "yes")) %>%
  pull("rep_rate")
nr_resp = nr_mail * rep_rate 
mail_cost = 1 * nr_mail 
profit = 9 * nr_resp - mail_cost
ROME = profit / mail_cost
```

Based on independent RFM the number of customers BBB should mail is `r format_nr(nr_mail, dec = 0)` (`r format_nr(perc_mail, perc = TRUE)`). The response rate for the selected customers is predicted to be `r format_nr(rep_rate, perc = TRUE)` or `r format_nr(nr_resp, dec = 0)` buyers. The expected profit is `r format_nr(profit,"$", dec = 0)`. The mailing cost is estimated to be `r format_nr(mail_cost, "$", dec = 0)` with a ROME of `r format_nr(ROME, perc = TRUE)`  


## Repsonse rate with sequential RFM

Create the `mailto_sq` variable for sequential RFM

    mailto_sq = mean(buyer == "yes") > 1/9

```{r}
## mail TRUE or FALSE
bbb <- group_by(bbb, rfm_sq) %>% 
  mutate(mailto_sq = (sum(training == 1 & buyer == "yes") / sum(training == 1)) > breakeven) %>% 
  ungroup()

## also calculate response rate per group as an illustration
bbb <- group_by(bbb, rfm_sq) %>% 
  mutate(rfm_sq_resp = sum(training == 1 & buyer == "yes") / sum(training == 1)) %>% 
  ungroup()
```

```{r}
result <- pivotr(
  bbb, 
  cvars = c("mailto_iq", "training"), 
  fun = "n_obs", 
  normalize = "column"
)
summary(result, dec = 2, perc = TRUE)
```

```{r}
result <- explore(
  bbb, 
  vars = "buyer", 
  byvar = c("training", "mailto_iq"), 
  fun = c("n_obs", "mean")
)
summary(result, dec = 4)
```

## Calculate profits and ROME

```{r}
dat <- bbb
perc_mail = mean(dat$mailto_sq)
nr_mail = 500000 * perc_mail
rep_rate <- filter(dat, mailto_sq == TRUE) %>% 
  summarize(rep_rate = mean(buyer == "yes")) %>%
  pull("rep_rate")
nr_resp = nr_mail * rep_rate 
mail_cost_sq = 1 * nr_mail 
profit_sq = 9 * nr_resp - mail_cost_sq
ROME_sq = profit_sq / mail_cost_sq
```

Based on sequential RFM the number of customers BBB should mail is `r format_nr(nr_mail, dec = 0)` (`r format_nr(perc_mail, perc = TRUE)`). The response rate for the selected customers is predicted to be `r format_nr(rep_rate, perc = TRUE)` or `r format_nr(nr_resp, dec = 0)` buyers. The expected profit is `r format_nr(profit_sq,"$", dec = 0)`. The mailing cost is estimated to be `r format_nr(mail_cost_sq, "$", dec = 0)` with a ROME of `r format_nr(ROME_sq, perc = TRUE)`.

Compare this to the main results from independent RFM. The expected profit is `r format_nr(profit,"$", dec = 0)`. The mailing cost is estimated to be `r format_nr(mail_cost, "$", dec = 0)` with a ROME of `r format_nr(ROME, perc = TRUE)`  

## Confirming the break-even response rate

If we select the predicted response rate for both of the rfm indices (i.e., `rfm_iq_resp` and `rfm_iq_resp`), and select a `profit` and `ROME` plot in `Model > Evaluate Classification` you should see the plots below. A visual inspection suggests that profits will be maximized if we target the top 35% of customers (approximately). You already calculated the exact percentages above. You should find that the number you calculated is very similar to the numbers highlighted green in the `performance.xls` file on Canvas.

```{r fig.width = 6, fig.height = 4, dpi = 244}
result <- evalbin(
  bbb, 
  pred = c("rfm_iq_resp", "rfm_sq_resp"), 
  rvar = "buyer", 
  lev = "yes", 
  qnt = 50, 
  margin = 9, 
  train = "Test", 
  data_filter = "training == 1"
)
summary(result, prn = FALSE)
plot(result, plots = "profit", custom = FALSE)
```

## Using the RFM index in a logistic regression


```{r fig.width = 7, fig.height = 5.38, dpi = 200}
result <- logistic(
  bbb, 
  rvar = "buyer", 
  evar = "rfm_sq", 
  lev = "yes",
  data_filter = "training == 1"
)
summary(result)
pred <- predict(result, pred_data = bbb)
print(pred, n = 10)
bbb <- store(bbb, pred, name = "predict_logit")
```

```{r fig.width = 10.5, fig.height = 4.67, dpi = 200}
visualize(
  bbb, 
  xvar = "rfm_sq", 
  yvar = "predict_logit", 
  type = "bar", 
  labs = list(
    y = "Predicted purchase probability", 
    x = "Logistic regression with Sq. RFM"
  ),
  data_filter = "training == 0",
  custom = TRUE
) +
  geom_hline(aes(yintercept = breakeven)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

