---
title: "DOE: Review"
output: html_document
---

```{r setup, include=FALSE}
## initial settings
knitr::opts_chunk$set(comment=NA, echo=FALSE, error = TRUE, cache=FALSE, message=FALSE, warning=FALSE)

## width to use when printing tables etc.
options(width = 250)

## make all required libraries available by loading radiant package if needed
if (!"package:radiant" %in% search()) library(radiant)

## finding the path to the class dropbox folder
fp <- file.path(find_dropbox(),"MGTA455-2017")
bizware <- readxl::read_excel(file.path(fp, "data/bizware.xlsx")) %>% na.omit
```

## Experimental design and partial factorials

### Generate a partial factorial design using information from the response table shown.

The smallest appropriate partial factorial has 8 trials (i.e., the solution with 8 trials has a perfect D-efficiency score equal to 1, no correlation between the variables (i.e., a determinant of 1), and is balanced (i.e., `trial` and `gift` both occur in 4 trials, `speed` and `power` both occur in 4 trials, and `$150`, `$160`, `$170`, and `$180` each occur in 2 trials).   

```{r}
fcts <- "price; $150; $160; $170; $180
message; speed; power
promotion; trial; gift"
result <- doe(fcts, seed = 172110)
summary(result)
```

Recall that multiple partial factorials may exist that "solve" the experimental design problem.

## Estimate a logistic regression based on the response table shown and predict response for all profiles

Open the `bizware.xlsx` file. After copy-and-pasting the data into Radiant we first need to create a _positive_ and _negative response_ variables. Then we can `gather` the data and estimate the logistic regression using (frequency) weights.

```{r}
## create new variable(s)
bizware <- mutate(bizware, yes = 2000 * response, no = 2000 - yes)

## Gather columns
bizware_gathered <- gather(bizware, resp, freq, yes, no, factor_key = TRUE)
```

The results from the logistic regression models are given below.

```{r}
result <- logistic(bizware_gathered, rvar = "resp", evar = c("price", "message", "promotion"), lev = "yes", wts = "freq")
summary(result)
```

The easiest way to generate predictions for all possible profiles (trials) if to use `expand.grid` and select the newly created dataset for prediction.

```{r}
## expanding data
bizware_gathered_expand <- expand.grid(level_list(bizware_gathered, price, message, promotion))
```

```{r}
result <- logistic(bizware_gathered, rvar = "resp", evar = c("price", "message", "promotion"), lev = "yes", wts = "freq")
summary(result)
pred <- predict(result, pred_data = "bizware_gathered_expand", prn = 16)
store(pred, data = "bizware_gathered_expand", name = "pred_logit")
print(bizware_gathered_expand)
```

See also the following <a href="https://www.youtube.com/watch?v=bdFDAu18O20" target="_blank">tutorial video</a>.

